Causal graph analysis of COVID-19 observational data in German districts reveals effects of determining factors on reported case numbers

### 2021-03-23 manuscript/v5_submission_revision/replyreviewer.Rmd

![Diagnostic plots: independence of observations, temporal.](figuresreviewer/independence_obs_resids_vs_time.png)

![Diagnostic plot: independence of observations, districts.](figuresreviewer/independence_id_vs_resid.png)
\newpage

**2) There are no tests for multi-collinearity that are presented, which presents significant concerns for over-fitting the model. Temperature and humidity are an example here. So is the mobility data.**

5. This an important remark which led to some major revisions on the analysis pipeline and the paper. We assessed variance inflation factors (VIFs) for the full model. We observed high VIFs for mobility variables and intervention variables. We decided to transform the original mobility variables by PCA and used the set of these principal components instead of the original mobility variables. Furthermore, we employed ridge regression as a way to further deal with multicollinearity and overfitting. We want to add that causal graph analysis with regression is more robust to multicollinearity than traditional regression approaches (see Schisterman et al 2017), and that in the presence of correlated variables, one might miss important effects, but the effects found are still valid. To find the most appropriate hyper-parameter for the ridge regression, we employed cross-validation on subsets of the 401 districts. Due to employing ridge regression for the effect estimates, we cannot report meaningful confidence intervals/p-values. Furthermore, we used cross-validation for identifying the most appropriate lag parameter, which was subsequently changed from 8 to 5 days in the manuscript.

### 2021-03-23 manuscript/v4_medrxiv_revision/causal_model.Rmd

```{r dag-covid-19, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="DAG of determinants of reported COVID-19 cases on the district level. Unobserved variables are light gray, variables marked with an asterisk (*) are confounded by weekday/holiday.", fig.height=5}
knitr::include_graphics(rep('figures/f_full_covid19_dag.eps'))
```

>>>>

```{r dag-covid-19, echo=FALSE, message=FALSE, warning=FALSE, fig.cap="DAG of determinants of reported COVID-19 cases on the district level. Unobserved variables are light gray, variables marked with an asterisk (*) are confounded by weekday/holiday.", fig.height=5}
knitr::include_graphics(rep('figures/f_full_covid19_dag_klein.eps'))
```
